---
title: "wk4_homework_casa5"
author: "a-saveleva"
date: "2024-10-24"
output: html_document
---
```{r include=FALSE}
library(tidyverse)
library(dplyr)
library(sf)
library(tmap)
library(tmaptools)
library(countrycode)
library(remotes)
```

Reading the csv file, and take the following columns:
- country
- gii_2010
- gii_2019

Drop the NA values.

```{r setup, include=FALSE}
gii <- read_csv('HDR23-24_Composite_indices_complete_time_series.csv', locale = locale(encoding = "latin1") %>%
  select(.,
        c("iso3","country","gii_2010", "gii_2019")) %>%
  drop_na() %>%
  filter(!str_detect(iso3, "^ZZ.*"))
```
Changing the ISO code:
```{r}
gii <- gii %>%
  mutate(iso2c = countrycode(iso3, "iso3c", "iso2c"))
```
Create a difference column:
```{r}
gii <- gii %>%
  mutate(.,
         gii_rel_change = (gii_2019 / gii_2010) - 1
  )
```
Read the shapefile.
!!!! Replace the name of the file if you downloaded geojson or other file type !!!!!
```{r include=FALSE}
shapes <- st_read(
  "World_Countries_(Generalized)_2402777631520798174/World_Countries_Generalized.shp"
) %>%
  st_simplify(., dTolerance = 300, preserveTopology = TRUE)
```
Merge csv into the shapefile:
```{r}
gii_shapes <- full_join(shapes, gii, by = c("ISO" = "iso2c"))
```
Plot:
```{r}
tmap_mode = "plot"
tm_shape(gii_shapes) +
  tm_polygons("gii_rel_change",
              palette="plasma",
              midpoint=0,
              title = "Change in global gender inequality between 2010 and 2019")
```

The map shows that most countries had seen a decrease of index of gender inequality (gii). A low GII value indicates low inequality between women and men, and vice-versa.

The country with the highest decrease is:

```{r echo=FALSE}
min_value <- gii[(which.min(gii$gii_rel_change)), ]
print(min_value[,2])
```

The country with the highest increase is:

```{r echo=FALSE}
max_value <- gii[(which.max(gii$gii_rel_change)), ]
print(max_value[,2])
```

The countries that had increase in gender inequality are:


```{r echo=FALSE}
print(gii %>% filter(gii_rel_change > 0)) %>% arrange(desc(gii_rel_change))
```


